Article Title

### 2013-12-24 plos.tex
minor edits of Methods.

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation.  The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.

In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.

@@

Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits.  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

@@

\paragraph{Additional information in neural correlation matrices}
Yet, a complete correlation matrix provides more information than the equivalent number of isolated pairwise correlations. In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}. In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Such approximations are particularly suitable for capturing shared fluctuations across the recorded population. Low-rank components can be extracted using principal component or factor analysis. The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.

Alternatively, the correlation matrix can be expressed through the partial pairwise correlations. The partial pairwise correlation is the linear correlation remaining after accounting for correlations with all the other neurons. Partial correlations carry particular significance when the relationship between the variables is approximately linear. In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent.

Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, networks of partial correlations (sometimes called \emph{association networks}) can be used to infer interactions between the components of a complex system.  This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies. A multivariate normal distribution specified by a weighted graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.

@@

The true correlation matrix is the normalized version of the true covariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  The usual estimator of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$,  using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.  Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate. Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{ 01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Model selection}
Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit may provide the best descriptor of the low-dimensional representation of the correlation structure of the system of interest. Its structure may then be analyzed and related to the circuit's anatomical organization.  
\paragraph{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' (\figref{02}{row 1}).

@@

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $LL^\T + \Psi$, where $L$ is the $p\times d$ matrix of \emph{factor loadings}, where $d$ is the number of factors, and $\Psi$ is the diagonal matrix of independent variances. The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (LL^\T + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).  This assumption is enforced by setting to zero an optimal subset of the off-diagonal elements of the precision matrix.  Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}). The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

@@

In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5). For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth. Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.

@@

\paragraph{Covariance estimation in neural data}
We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus. Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}. This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis (See Methods). The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}). Overall, 31 imaged sites from 24 animals were included in the study.

@@

\paragraph{Correlation structure and circuit architecture}
Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites. By design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units. If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations. In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}). At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}). However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}). The partial correlation was decomposed into sparse and low-rank components (\figref{ 05}{C}). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations. The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}). The low-rank component had rank 17.

@@

In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients at a level deemed significant. A network was formed by connecting neurons whose total correlations exceed this threshold \cite{Golshani:2009,Malmersjo:2013}. There was fairly little overlap between the network of interactions revealed by such thresholding and those revealed by the sparse+latent estimator (\figref{05}{F}). When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (\figref{05}{F, H, and I}). In particular, many low sample correlations translated into negative interactions in the regularized estimate. Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient. Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.

@@

We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}). The average partial correlations fell more rapidly with difference in preferred orientation (\figref{fig:07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).

@@

\section*{Discussion}
\paragraph{Functional and anatomical connectivity}
Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.  This line of inquiry requires expressive descriptions of the observed activity of neural circuits.

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation[1].  The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.

In this study, we sought most statistically efficient estimators of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 200 $\mu$m wide and 100 $\mu$m deep (150--350 cells) in mouse visual cortex.  We hypothesized that in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of partial correlations between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator based on such approximations outperformed other regularized estimators. Mirroring previously described patterns of synaptic connectivity, the density of positive partial correlations decreased rapidly with both physical distances and preferred orientation differences whereas negative partial correlations were less selective.  The inferred functional connectivity obtained using these methods can generate hypotheses about synaptic architecture that can guide future experimental work.[2]

@@

Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are among the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006, Zohary:1994, Kohn:2005, Bair:2001, Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of trial-to-trial response variability between pairs of neurons, have a profound impact on stimulus coding \cite{Zohary:1994, Abbott:1999, Sompolinsky:2001, Averbeck:2006, Josic:2009, Berens:2011, Ecker:2011}. In addition, noise correlations and correlations in spontaneous activity have been hypothesized to reflect key features of functional connectivity in neural circuits[3].  This interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001, Arieli:1995, Chiu:2002, Kenet:2003, Kohn:2005, Cohen:2008, Cohen:2009, Ko:2011, Smith:2013b[4]}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009, Kohn:2009, Ecker:2010, Renart:2010}, and others.

@@

\paragraph{Additional information in neural correlation matrices}
Yet, a complete [5]correlation matrix provides more information than the equivalent number of isolated pairwise correlations. In early studies, the effects of correlations were extrapolated from isolated pairs to entire populations through simulation and theoretical analysis \cite{Shadlen:1998, Zohary:1994}. In contrast, modern multineuronal recordings from increasingly large and dense populations of highly interconnected neurons allow estimation of various low-dimensional representations of the entire correlation matrix, which can enable inquiry into the nature of underlying physiological mechanisms and computational principles.

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Such approximations are particularly suitable for capturing shared fluctuations across the recorded population.[6] Low-rank components can be extracted using principal component or factor analysis. The resulting structure of  neural correlations has been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009[7]}. [8]

Alternatively, the correlation matrix can be expressed through the partial pairwise correlations. The partial pairwise correlation is the linear correlation remaining[9] after accounting for correlations with all the other neurons. Partial correlations carry particular significance when the relationship between the variables is approximately linear. In this case, small pairwise partial correlations indicate that the two variables are nearly conditionally independent[10].

Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, networks of partial correlations (sometimes called \emph{association networks}) can be used to infer interactions between the components of a complex system.  This approach has been used to uncover the structure of gene interaction networks \cite{Schafer:2005,Peng:2009} and the functional connectivity between brain regions from fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In \emph{elliptical distributions} [11](those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial correlations express conditional dependencies. A multivariate normal distribution specified by a weighted graph[12] of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With departure from linearity, the correspondence between conditional dependence and partial correlations diminishes and can break down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006, Tkacik:2006}.[13]

@@

The true correlation matrix is the normalized version of the true covariance matrix defined as
\begin{equation}\label{eq:true-covariance}
    \Sigma = \E{(x-\mu)(x-\mu)^\T},\quad \mu = \E{x}
\end{equation}
where $\E{\cdot}$ denotes expectation; $x$ is the $p\times 1$ vector of real-valued instantaneous firing rates in bins of duration $\Delta t$; and $\mu$ is the vector of mean firing rates. In the case of noise correlations, $\mu$ depends on the stimulus condition.  The usual estimator [14]of the covariance matrix is the \emph{sample covariance matrix} $C_{\sf 0}$ computed from the empirical sample of observations $x(t),\; t=n(1,\ldots,n)$ as
\begin{equation}
    C_{\sf 0} = \frac 1 \nu \sum\limits_{t=1}^n (x(t)-\bar x)(x(t)-\bar x)^\T,\quad \bar x= \frac 1 n \sum\limits_{t=1}^n x(t)
\end{equation}
where $\nu$ is the number of degrees of freedom per neuron in the sample ($\nu=n-1$ if observations are independent). In this study, we estimate the true mean, $\mu$,  using the sample mean $\bar x$, but seek a better estimate of the covariance matrix than $C_{\sf 0}$.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The sample estimate $C_{\sf 0}$ is unbiased ($\E{C_0}=\Sigma$) but, on average, falls far from $\Sigma$ due to its sensitivity to sampling noise.  Low-dimensional estimates are typically less susceptible to sampling noise but can introduce biases.  Regularization strikes a favorable balance between bias and variability, and can produce \emph{some} improvement even with an arbitrary target estimate  [15]Yet, when the target estimate captures important features of the true covariance matrix with few terms, a regularized estimate introduces minimal bias and outperforms other estimators.

\paragraph{Illustration of estimation of a neural correlation matrix}[16]
To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\begin{figure}[!ht]    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Illustration of regularized estimation of partial correlations.}
        {\bf A}. The sample correlation matrix of unprocessed somatic calcium signals from a population of cells in mouse visual cortex.
        The outlined square fragment is magnified in {\bf B}.
        {\bf C}. The same fragment of another estimate of the correlation matrix regularized to yield sparse partial correlations.
        Corresponding fragments of partial correlations matrices of the unregularized and regularized estimated are shown in {\bf D} and {\bf E}, respectively.
    }
    \label{fig:01}}
    {\includegraphics[width=8.3cm]{./figures/Figure01.pdf}}
\end{figure}
\paragraph{Model selection}
Covariance selection is but one of many conceivable low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant regularities in the activity of the circuit, its so-called `functional connectivity', tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits have different types of regularities in their activity, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit may provide [17]the best descriptor of the low-dimensional representation of the correlation structure of the system of interest. Its structure may then be analyzed and related to the circuit's anatomical organization.  
\paragraph{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations. [18]

\section*{Results}
% Results and Discussion can be combined.
\paragraph{Covariance estimation}
We considered four regularized estimators based on distinct families of low-dimensional target estimates: `independent', `latent factors', `sparse interactions', and `sparse+latent' [19](\figref{02}{row 1}).

@@

The target of the second regularized estimator $C_{\sf factor}$, is the factor model $LL^\T + \Psi$, where $L$ is the $p\times d$ matrix of \emph{factor loadings}, where $d$ is the number of factors, and $\Psi$ is the diagonal matrix of independent variances. The estimator, 
\begin{equation}\label{eq:c-factor}
C_{\sf factor} = (1-\lambda) C_{\sf 0} + \lambda (LL^\T + \Psi),
\end{equation}
has two hyperparameters: the number of factors $d$ and shrinkage intensity $\lambda$. The target estimate $LL^\T + \Psi$ has the structure that arises when correlated fluctuations in population activity are linearly driven by a small number [20]of latent factors that affect many cells while direct interactions between cells are insignificant (\figref{02}{B}).

The third estimator $C_{\sf sparse}$ is based on the assumption that all correlations result from direct linear associations between a sparse set of pairs of observed cells (\figref{02}{C}).  This assumption is enforced by setting to zero an optimal[21] subset of the off-diagonal elements of the precision matrix.  Therefore, $C_{\sf sparse}$ is biased towards correlation structures that arise in networks of linearly interacting pairs of neurons (\figref{02}{C}). The estimator takes the form
\begin{equation}\label{eq:c-sparse}
C_{\sf sparse} = S^{-1},
\end{equation}
where $S$ is a sparse matrix with a large fraction of zeros off the diagonal. The estimate has one hyperparameter that determines the sparsity (fraction of off-diagonal zeros) of $S$.

@@

In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5). For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth. Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples taken for each ground truth and for each sample size.[22]

@@

\paragraph{Covariance estimation in neural data}
We recorded calcium activity in dense populations of neurons in the supragranular layers in primary visual cortex of anesthetized mice using fast random-access 3D scanning two-photon microscopy \cite{Reddy:2005, Katona:2012, Cotton:2013}. We presented 300 repetitions of full-field drifting gratings with two directions of motion or 150 repetitions with five directions (\figref{03}{A}) on one side of the visual field of anesthetized mice. Groups of cells loaded with a calcium-sensitive fluorescent dye were imaged and localized in 3D space (\figref{03}{B and E} in the visual cortex on the contralateral side to the stimulus. Using acousto-optic deflectors (AOD) to steer the laser in 3D,  we recorded the somatic calcium activity from the located cells with concurrent motion detection \cite{Cotton:2013}.  [23]This technique allowed fast sampling (100--150 Hz) from large numbers (150--350) of cells in a small volume of cortical tissue ($200\times200\times100$ $\mu$m$^3$) in layers 2/3 and 4 (\figref{03}{C}). After downsampling the signal to 20 Hz, relative firing rates were inferred using sparse nonnegative deconvolution \cite{Vogelstein:2010} (\figref{03}{C and D}). Only cells that produced detectable calcium activity were included in the analysis[24] (See Methods).[25] The average stimulus response was subtracted from each trial; the remaining signals were further downsampled into 150 ms bins to compute the noise correlation matrix (\figref{ 03}{F and G}). The durations of recordings dedicated to estimating the noise correlations ranged between 15--20 minutes.  Besides the high-repetition stimulus protocol used to estimate noise correlations, an orientation tuning stimulus protocol was used to map the orientation tuning of each cell (\figref{03}{E}). Overall, 31 imaged sites from 24 animals were included in the study.

@@

\paragraph{Correlation structure and circuit architecture}
Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the low-dimensional structure revealed by this estimator for individual imaged sites. By design, the sparse+latent estimator finds the optimal[26] balance between a sparse network of partial correlations and shared common latent units. If this estimate dominates, we can hypothesize that it is better reveals underlying physiological interactions than the usual correlations[27]. [28]In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}). At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}). However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}). The partial correlation was decomposed into sparse and low-rank components (\figref{05}{C}). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations. [29]The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}). The low-rank component had rank 17.[30]

@@

In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients at a level deemed significant. A network was formed by connecting neurons whose total correlations exceed this threshold \cite{Golshani:2009,Malmersjo:2013}. There was fairly little overlap between the network of interactions revealed by such thresholding and those revealed by the sparse+latent estimator (\figref{05}{F}). When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (\figref{05}{F, H, and I}). In particular, many low sample correlations translated into negative interactions in the regularized estimate. Indeed, the absence of correlations between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient. [31]Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.

@@

We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}). The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).[32]

@@

\section*{Discussion}
\paragraph{Functional and anatomical connectivity}
Systems neuroscience is the quest to discover principles whereby anatomical organization of neural circuits gives rise to their function.  This line of inquiry requires expressive descriptions of the observed activity of neural circuits.[33]
